---
title: "project_Hyperboria"
author: "Rémy Cerda and Valentin Lorentz"
date: "29 novembre 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library("ggplot2")
library("dplyr")
```

# Introduction

This project is about analysing latency and reachability on a mesh network, sometimes using a partial knowledge of its topology.
We collected data on the Hyperboria network, which is a mesh network deployed both at local (over Ethernet) and global (over UDP) levels.
Hyperboria has the advantage of making its list of nodes easily discoverable (it uses source routing), so we can get a list of most of the nodes of the network easily.

The database of pings it created by running 3 pings to each node, every 10 minutes, and storing the latency.

# Producing data

On a computer connected to Hyperboria (and with `fc00::/8` routed to it), create the file “scan_data.csv” with this content:

```
timestamp,round_start_timestamp,node,seq,nb_bytes,ttl,latency
```

Then, run:

* `python3 ping_scan.py` wait for the results, press Ctrl-C. This will produce `scan_data.csv`.
* `gzip scan_data.csv` (the result is in this git repo)
* `cat scan_data.csv.gz | gunzip | sed "s/ ms//" | gzip > scan_data_fixed.csv.gz`

And:

* `wget https://www.fc00.org/static/graph.json` (the result is in this git repo)
* `python3 compute_distances.py`. This will produce `node_distances.csv`.



# Load data

## Data structure

Each line is a ping sent to an Hyperboria node, using Debian's `ping`.

Columns are:

* timestamp: timestamp the ping was sent, as reported by ping.
* round_start_timestamp: timestamp the `ping` tool was invoked at.
* node: IPv6 address of the pinged node.
* seq: sequence number, as reported by ping. There is a bug with my script: for nodes that can be pinged, the seq is from 2 to 4, and for nodes which could not be reach it is from 0 to 2. The former should be fixed before analysing data.
* nb_bytes: number of bytes of the ICMP PING, as reported by ping
* ttl: ICMP TTL, as reported by ping
* latency: latency reported by ping.

```{r}
pings = read.csv("scan_data_fixed.csv.gz")
head(pings)
summary(pings)
```

## Fixes

Compensate offset bug of the ping collection script.

```{r}
lines_with_seq_offset = !is.na(pings$latency)
pings$seq[lines_with_seq_offset] = pings$seq[lines_with_seq_offset]-2
summary(pings$seq)
```

## Distances database

The distance database is computed from the graph hosted at fc00.org. This graph is obtained from various probes on the network submitting their view of the network to fc00.org, which aggregates results.

Load the database:

```{r}
distances = read.csv("node_distances.csv")
```

Add distance column to the ping data:

```{r}
data = merge(pings, distances, by="node", all.x=TRUE)
head(data)
summary(data)
```

# Analysis

## Latency distribution

First, let us have a look to the repartition of latency:
```{r}
summary(data$latency)
ggplot(data=data, aes(x=latency)) + xlim(0, 2000) + geom_histogram(binwidth=100);
```

## Latency and distance

Let us draw histograms of latency for each distance:
```{r}
ggplot(data=data, aes(x=latency)) + xlim(0, 2000) + geom_histogram(binwidth=50) + facet_grid(distance ~ .);
```

The graph is not really readable. Further analysis can be led using following boxplots:
```{r}
ggplot(data=data, aes(x=factor(distance), y=latency)) + geom_boxplot();
```

And for readability, a log-scaled version:
```{r}
ggplot(data=data, aes(x=factor(distance), y=latency)) + geom_boxplot() + scale_y_log10();
```

## Lag graph of latency

First, compute the difference of latency for each node, between sequences of 3 pings:

```{r}
aggregated_data = data %>% group_by(round_start_timestamp, node) %>% summarize(latency=mean(latency), distance=min(distance)) %>% filter(!is.na(latency)) %>% as.data.frame();
summary(aggregated_data$latency)

head(aggregated_data)
df_lag = aggregated_data %>% group_by(node) %>% mutate(old_latency=dplyr::lag(latency)) %>% mutate(latency_change=latency - old_latency) %>% mutate(relative_latency_change=latency_change/(old_latency+latency)) %>% as.data.frame()
head(df_lag)
```

```{r}
#ggplot(data=df_lag, aes(x=old_latency, y=latency)) + xlim(0, 2500) + ylim(0, 2500) + geom_point()
```

We can observe the distribution of latency changes:

```{r}
ggplot(data=df_lag, aes(x=latency_change)) + xlim(-1000, 1000) + geom_histogram(bins=100)
summary(df_lag$latency_change);
```

Scaled to the latency of each node:

```{r}
ggplot(data=df_lag, aes(x=relative_latency_change)) + geom_histogram(bins=100)
#ggplot(data=df_lag, aes(x=latency, y=relative_latency_change)) + geom_point()
summary(df_lag$relative_latency_change)
```

Finally, we can split this distribution over the different distances, which all seem to have a similar distribution:

```{r}
ggplot(data=df_lag, aes(x=relative_latency_change)) + xlim(-2,2) + geom_histogram(bins=50) + facet_grid(distance ~ ., scale="free_y");
```


## Distribution of number of replies to ping from a node

Here, we are not interested in the latency, but whether or not a node replies.

First, compute the number of rounds in which each node responded (to the third ping), discarding nodes which never replies:

```{r}
df_counts = data %>% filter(seq == 2) %>% filter(!is.na(latency)) %>% group_by(node) %>% summarize(nb_replies=n(), distance=mean(distance)) %>% as.data.frame();
summary(df_counts)
head(df_counts)
```

Although we cannot see any apparent pattern on global statistics...

```{r}
ggplot(data=df_counts, aes(x=nb_replies/max(nb_replies))) + geom_histogram(bins=50);
```

... we can see a mostly-linear decrease of the ratio replies when the distance increases.
Note that NA is for nodes which are not on the graph of fc00 -- usually because none of the probes know where on the graph they are, which is either caused by them being far for the probe / hard to reach / not very reliable.

We can also note that nodes at distance 1 are highly reliable -- a potential bias to this is that they are direct peers of the source node (Hydrogen), which I chose myself for their reliability. Nodes at distance >= are more evenly distributed, because Hydrogen's peers are far more open on their peering policy.

```{r}
ggplot(data=df_counts, aes(x=factor(distance), y=nb_replies/max(nb_replies))) + geom_boxplot();
```
