---
title: "project_Hyperboria"
author: "Rémy CERDA and Valentin Lorentz"
date: "29 novembre 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load data

```{r, include=FALSE}
library(ggplot2)
```

## Data structure

Each line is a ping sent to an Hyperboria node, using Debian's `ping`.

Columns are:

* timestamp: timestamp the ping was sent, as reported by ping.
* round_start_timestamp: timestamp the `ping` tool was invoked at.
* node: IPv6 address of the pinged node.
* seq: sequence number, as reported by ping. There is a bug with my script: for nodes that can be pinged, the seq is from 2 to 4, and for nodes which could not be reach it is from 0 to 2. The former should be fixed before analysing data.
* nb_bytes: number of bytes of the ICMP PING, as reported by ping
* ttl: ICMP TTL, as reported by ping
* latency: latency reported by ping.

```{r}
data = read.csv("scan_data.csv.gz")
head(data)
summary(data)
```

## Fixes

Compensate offset bug of the ping collection script.

```{r}
lines_with_seq_offset = !is.na(data$latency)
data$seq[lines_with_seq_offset] = data$seq[lines_with_seq_offset]-2
summary(data$seq)
```

# Analysis

First, let us have a look to the repartition of latency:
```{r}
hist(c(data$latency))
```
